Technique for \earlywarning\ detection of compact binary coalescence and its implications for multi-messenger astronomy

### 2011-06-14 introduction.tex
Added LSST comparison
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1075 ccdd05d5-56c1-461a-bf98-85f6876523b4

\subsection{Rapid \EM{} followup}

@@

Telescopic followup requires estimating the location of the \GW{} source. The
localization uncertainty is dominated by uncertainty in the time of arrival of
the \GW{}s, which is determined by the signal's effective bandwidth and \SNR{}
\citep{Fairhurst2009}. Unfortunately, for a \GW{} signal with an \SNR{} of
blah, the 90\% confidence area is at least blah deg$^2$ (blah\% of the sky) at
10~s before merger and blah deg$^2$ (blah\% of the sky) at 1~s before merger.
If we can take galactic luminosities to accurately trace mass and thus proxy
for the \CBC{} event rate, it is possible to reduce the localization
uncertainty using a galaxy catalog as a prior probability distribution
\citep{galaxy-catalog}, along with the sky location and luminosity distance
estimate from the \GW{} signal, as was done in S6/VSR3. At Advanced \LIGO{}
\textsc{ns}--\textsc{ns} horizon distances of $\sim$400~Mpc, the galaxy catalog
is fairly isotropic and thus not as useful for downselecting pointings, but
sources that are detectable before merger are likely to be exceptionally
nearby. Thus there will be relatively few galaxies to consider for a given
candidate and the galaxy catalog is likely to be sufficiently complete. This
may allow us to shrink the 90\% confidence area substantially. Even so, it will
be desirable to mobilize a large number of heterogeneous telescopes in order to
guarantee imaging the source at the instant of merger. Because the sky
localization will improve rapidly down to blah deg$^2$ at merger, by the time
that the entire \CBC{} event is over it will be possible to downselect the
images that have already been taken as well as repoint telescopes to more
likely locations.

\subsection{Low-latency \CBC{} search algorithms}

>>>>

\subsection{Early-warning \EM{} followup}

@@

Telescopic followup requires estimating the location of the \GW{} source. The
localization uncertainty is dominated by uncertainty in the time of arrival of
the \GW{}s, which is determined by the signal's effective bandwidth and \SNR{}
\citep{Fairhurst2009}. Unfortunately, for a \GW{} signal with an \SNR{} of
blah, the 90\% confidence area is at least blah deg$^2$ (blah\% of the sky) at
10~s before merger and blah deg$^2$ (blah\% of the sky) at 1~s before merger.
It is infeasible to search this large an area for a prompt counterpart. For
reference, \textsc{lsst} should see first light around the same time as
Advanced \LIGO{} and will have an unparalleled field of view of 9.6 deg$^2$
\citep{2008arXiv0805.2366I}.

If we can take galactic luminosities to accurately trace mass and thus proxy
for the \CBC{} event rate, it is possible to reduce the localization
uncertainty using a galaxy catalog as a prior probability distribution
\citep{galaxy-catalog}, along with the sky location and luminosity distance
estimate from the \GW{} signal, as was done in S6/VSR3. At Advanced \LIGO{}
\textsc{ns}--\textsc{ns} horizon distances of $\sim$400~Mpc, the galaxy catalog
is fairly isotropic and thus not as useful for downselecting pointings, but
sources that are detectable before merger are likely to be exceptionally
nearby. Thus there will be relatively few galaxies to consider for a given
early-warning candidate and the galaxy catalog is likely to be sufficiently
complete. This should allow us to shrink the 90\% confidence area substantially.
Even so, it will be desirable to mobilize a large number of heterogeneous
telescopes in order to guarantee imaging this exceptional event at the instant
of merger. Because the sky localization will improve rapidly down to blah
deg$^2$ at merger, by the time that the entire \CBC{} event is over it will be
possible to send a second alert to downselect the images that have already been
taken as well as repoint telescopes to more likely locations.

\subsection{Latency in \CBC{} search algorithms}
